---
title: "Example visuals"
author: "Adeline Lo"
date: "6/17/2020"
output: 
  html_document: 
    pandoc_args: ["--lua-filter=color-text.lua"]
  pdf_document: 
    pandoc_args: ["--lua-filter=color-text.lua"]
    keep_tex: true
---

```{cat, engine.opts = list(file = "color-text.lua")}
Span = function(span)
  color = span.attributes['color']
  -- if no color attribute, return unchange
  if color == nil then return span end
  
  -- tranform to <span style="color: red;"></span>
  if FORMAT:match 'html' then
    -- remove color attributes
    span.attributes['color'] = nil
    -- use style attribute instead
    span.attributes['style'] = 'color: ' .. color .. ';'
    -- return full span element
    return span
  elseif FORMAT:match 'latex' then
    -- remove color attributes
    span.attributes['color'] = nil
    -- encapsulate in latex code
    table.insert(
      span.content, 1,
      pandoc.RawInline('latex', '\\textcolor{'..color..'}{')
    )
    table.insert(
      span.content,
      pandoc.RawInline('latex', '}')
    )
    -- returns only span content
    return span.content
  else
    -- for other format return unchanged
    return span
  end
end
```

```{r setup, echo=FALSE,eval=TRUE, warning = FALSE, message = FALSE}
library(tidyverse)
library(readr)
library(ggplot2)
library(countrycode)
library(paletteer) 
library(grid)
library(viridis)
library(scales) # to access break formatting functions
#grid stuff
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
library(ggExtra)
#tables
library(xtable)
#colors
plot_col<-viridis(n=5,alpha=1,begin=0,end=1,option="E")
plot_col2<-viridis(n=5,alpha=1,begin=0,end=1,option="D")
```


```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
tmp_file<-"../Data/Final results/table_covid_other.csv"
## Fig 1 data
Other   <- read_csv(paste(tmp_file)) %>% 
  #select(-Date,-X1) %>% 
  pivot_longer(4:11, 
               names_to = "YLL_type",
               values_to = "YLL") %>% 
  mutate(YLL_name = 
           case_when(YLL_type == "YLL_rates_covid" ~ "COVID",
                     YLL_type == "YLL_rates_heart" ~ "Heart",
                     YLL_type == "YLL_rates_drug" ~ "Drug",
                     YLL_type == "YLL_rates_transport" ~ "Transport",
                     YLL_type == "YLL_rates_flu_max" ~ "Flu (max)",
                     YLL_type == "YLL_rates_flu_med" ~ "Flu (med)",
                     YLL_type == "YLL_excess_ratio" ~ "Excess",
                     YLL_type == "YLL_projection_ratio" ~ "Projection"))

Base <- 
  Other %>% 
  pivot_wider(id_cols = c("Country"),
              names_from = "YLL_name", 
              values_from = "YLL") %>% 
  mutate(COVmean = mean(COVID))

## Create proportions: COVID/Other --> Other
Base$Heart<-Base$COVID/Base$Heart
Base$Drug<-Base$COVID/Base$Drug
Base$Transport<-Base$COVID/Base$Transport
Base$`Flu (max)`<-Base$COVID/Base$`Flu (max)`
Base$`Flu (med)`<-Base$COVID/Base$`Flu (med)`
Base$Excess<-1/Base$Excess
Base$Continent<-countrycode(as.character(Base$Country),origin='country.name',destination='continent')
Base$Continent[Base$Country=="England"]<-"Europe"
Base$Continent_col[Base$Continent=="Africa"]<-plot_col2[1]
Base$Continent_col[Base$Continent=="Americas"]<-plot_col2[2]
Base$Continent_col[Base$Continent=="Asia"]<-plot_col2[3]
Base$Continent_col[Base$Continent=="Europe"]<-plot_col2[4]
Base$Continent_col[Base$Continent=="Oceania"]<-plot_col2[5]
Base$Country<-factor(Base$Country,levels=Base$Country)
##Pull in days since first covid data -- need to update always to latest dates manually
days<-read.csv("../Data/days_since_first_covid_data.csv")
matched<-match(as.character(Base$Country),as.character(days$Country))
Base$Diff.Days<-days$Diff.Days[matched[which(!is.na(matched))]]
##
Base2<-Base
Base2<-Base2[order(Base2$`Flu (med)`,Base2$Transport,Base2$Heart,Base2$Excess,decreasing=FALSE),]
Base2$Country<-factor(as.character(Base2$Country),levels=as.character(Base2$Country))
Base2$Country_day<- paste(Base2$Country," (",Base2$Diff.Days,")",sep="") #variable with Country followed by day in parentheses
Base2$Country_day<-factor(as.character(Base2$Country_day),levels=as.character(Base2$Country_day))
all_countries<-sort(unique(Base$Country))

### Fig 2 data
gender<-readRDS("../Data/Final results/YLL_covid19_complete.rds")
gender<-gender[which(gender$Country!="Cameroon"),]#subset out NA, drop Cameroon (weirdly large)
gender$Continent<-countrycode(as.character(gender$Country),origin='country.name',destination='continent')
gender$Continent[gender$Country=="England"]<-"Europe"
gender$Continent_col[gender$Continent=="Africa"]<-plot_col2[1]
gender$Continent_col[gender$Continent=="Americas"]<-plot_col2[2]
gender$Continent_col[gender$Continent=="Asia"]<-plot_col2[3]
gender$Continent_col[gender$Continent=="Europe"]<-plot_col2[4]
gender$Continent_col[gender$Continent=="Oceania"]<-plot_col2[5]
gender<-gender[order(gender$Continent,gender$Country,decreasing=TRUE),]
gender$Country<-factor(gender$Country,levels=gender$Country)
gender$Rate<-gender$YLL_rates_male_to_female
gender2<-gender
gender2<-gender2[order(gender2$Rate,gender2$Country,decreasing=FALSE,na.last=FALSE),]
gender2$Country<-factor(as.character(gender2$Country),levels=as.character(gender2$Country))
gender2$Rate_color<-ifelse(gender2$Rate>1,plot_col[4],ifelse(gender2$Rate<1,plot_col[1],"white"))
#add global average to bottom so it's above Taiwan in the plot
#create population weights for weighted average
pop<-read.csv("../Data/Processed/pop_complete.csv")
pop<-data.frame(Country=names(tapply(pop$Total,pop$Country,sum)),Population=tapply(pop$Total,pop$Country,sum))
gender2$pop<-pop$Population[match(gender2$Country,pop$Country)]
global_gender<-data.frame(Country=c("","WEIGHTED AVERAGE","AVERAGE"),Date=rep(NA,3),YLL.b_abs=rep(NA,3)
                  ,YLL.b_rates=rep(NA,3),YLL.m_rates=rep(NA,3),YLL.f_rates=rep(NA,3),Deaths=rep(NA,3)
                  ,Deaths_m=rep(NA,3),Deaths_f=rep(NA,3),YLL_per_death_b=rep(NA,3)
                  ,YLL_per_death_males=rep(NA,3),YLL_per_death_females=rep(NA,3)
                  ,YLL_rates_male_to_female=rep(NA,3),Continent=rep(NA,3),Continent_col=rep(NA,3)
                  ,Rate=c(NA
                          ,sum(gender2$Deaths/sum(gender2$Deaths,na.rm=TRUE)*gender2$YLL_rates_male_to_female,na.rm=TRUE) #weighted
                          #,sum(gender2$Rate*gender2$pop,na.rm=TRUE)/sum(gender2$pop,na.rm=TRUE)
                          ,mean(gender2$YLL.m_rates,na.rm=TRUE)/mean(gender2$YLL.f_rates,na.rm=TRUE) #avg
                          #,mean(gender2$Rate,na.rm=TRUE)
                          ) #NA, weighted avg, avg
                  ,Rate_color=c(NA,plot_col[5],plot_col[5]), pop=rep(NA,3))
gender2<-rbind(gender2,global_gender)
gender2$Country<-factor(as.character(gender2$Country),levels=as.character(gender2$Country))
age<-readRDS("../Data/Final results/table_YLL_age_cut_offs.rds")
data<-data.frame(Country=as.character(rep(age$Country,each=3))
                 ,Age=rep(c("Under 55","55-75","75+"),length(unique(age$Country)))
                 ,Proportion=c(t(age[,2:4]))#row unlist data
)
data<-data %>% mutate(Age = fct_relevel(Age,"75+","55-75","Under 55"))
data$Continent<-countrycode(as.character(data$Country),origin='country.name',destination='continent')
data$Continent[data$Country=="England"]<-"Europe"
data$Continent_col[data$Continent=="Africa"]<-plot_col2[1]
data$Continent_col[data$Continent=="Americas"]<-plot_col2[2]
data$Continent_col[data$Continent=="Asia"]<-plot_col2[3]
data$Continent_col[data$Continent=="Europe"]<-plot_col2[4]
data$Continent_col[data$Continent=="Oceania"]<-plot_col2[5]
data<-data[order(data$Continent,data$Country,decreasing=TRUE),]
data$Country<-as.character(data$Country)
data$Country<-factor(data$Country,levels=unique(data$Country))
age<-data[order(data$Proportion,decreasing=FALSE),]
#age2$Age<-factor(age2$Age, levels = c("Under 55", "55-75", "75+"))
  #order by under 55 proportion countries
tmp<-subset(age,Age=="Under 55")
tmp<-tmp[order(tmp$Proportion,decreasing=FALSE),]
age$Country<-factor(as.character(age$Country),levels=as.character(tmp$Country))
    #add global average at the bottom so it translates to top (above Bangladesh)
global_age<-data.frame(Country=rep(c("","GLOBAL AVERAGE"),each=3),Age=rep(c("Under 55","55-75","75+"),2)
                  ,Proportion=c(NA,NA,NA,0.325,0.439,0.236),Continent=rep(NA,6),Continent_col=rep(NA,6)) #get proportions fm Weighted Proportions for Figure 2
age2<-rbind(age,global_age)

#for SI plots the first data YLL_covid19_complete, so just sync up days since into it
si1<-readRDS("../Data/Final results/YLL_covid19_complete.rds")
matched<-match(as.character(si1$Country),as.character(days$Country))
si1$Diff.Days<-days$Diff.Days[matched[which(!is.na(matched))]]
```
## Figure 1:

### Flu Panel

**Baseline plot: Flu**

```{r, echo=FALSE,eval=TRUE,warning = FALSE}
#Fig 1 csv data saved:
#write.csv(Base2,file="../Data/Final results/figure1_data.csv")
flu_base_plot <-ggplot(Base2, aes(x=`Flu (med)`, y=Country)) +
    geom_point(shape=15,color=plot_col2[2], size=3.5, alpha=1) +
    geom_point(aes(x=`Flu (max)`,y=Country),shape=15,color=plot_col2[3],size=3.5) +
    #relabel y-axis tick marks
    scale_y_discrete(labels=paste(c(as.character(Base2$Country_day)))) + 
    #legend
    geom_point(data=data.frame(x=c(5.3,5.3),y=c(5,3))
               ,aes(x=x,y=y),shape=c(15,15),color=c(plot_col2[2:3]),size=3.5) +
    ggplot2::annotate(geom="text",x=c(5.2,5.2),y=c(4,2),label=c("Median influenza year" ,"Maximum influenza year"),size=3,col=c("gray20")) +
    ggplot2::annotate(geom="text",x=(mean(Base2$`Flu (med)`)+0.5),y=42,label=paste(round(mean(Base2$`Flu (med)`),3)),size=3.5,col=plot_col2[2])+
    ggplot2::annotate(geom="text",x=(mean(Base2$`Flu (max)`)-0.6),y=42,label=paste(round(mean(Base2$`Flu (max)`),3)),size=3.5,col=plot_col2[3])+
    #average lines
    geom_vline(xintercept=mean(Base2$`Flu (med)`) ,linetype="solid", color=plot_col2[2], 
               size=1.25, alpha = 0.65)+
    geom_vline(xintercept=mean(Base2$`Flu (max)`), linetype="solid", color=plot_col2[3], 
               size=1.25, alpha = 0.65)+
    theme_light() + ylab("") + 
    #xlab("Ratio of years of life lost due to COVID-19 vs seasonal influenza") +
    xlab("A. Ratio of years of life lost due to COVID-19 vs seasonal influenza") +
    theme(panel.border = element_blank(), axis.ticks.y = element_blank()
          , axis.text.y=element_text(colour="gray20"))
flu_base_plot
#Fig1_flu.pdf -- without xlab A
```

### Transport Panel

**Baseline plot: Transport**


```{r, echo=FALSE,eval=TRUE,warning = FALSE}
transport_base_plot <-ggplot(Base2, aes(x=Transport, y=Country)) +
    geom_point(shape=17,color=plot_col2[1], size=3.5, alpha=1) +
    #relabel y-axis tick marks
    scale_y_discrete(labels=paste(c(as.character(Base2$Country_day)))) + 
    #average lines
    geom_vline(xintercept=mean(Base2$Transport) ,linetype="solid", color=plot_col2[1], 
               size=1.25, alpha = 0.65) +
    ggplot2::annotate(geom="text",x=(mean(Base2$Transport)+0.3),y=42,label=paste(round(mean(Base2$Transport),3)),size=3.5,col=plot_col2[1])+
    theme_light() + ylab("")+ 
    #xlab("Ratio of years of life lost due to COVID-19 vs transport") +
    xlab("B. Ratio of years of life lost due to COVID-19 vs transport") +
    theme(panel.border = element_blank(), axis.ticks.y = element_blank()
          , axis.text.y=element_text(colour="gray20"))
#transport_base_plot
#Fig1_transport.pdf -- without xlab B
```

### Heart Panel

**Baseline plot: Heart**

```{r, echo=FALSE,eval=TRUE,warning = FALSE}
heart_base_plot <-ggplot(Base2, aes(x=Heart, y=Country)) +
    geom_point(shape=16,color=plot_col2[4], size=3.5, alpha=1) +
    #relabel y-axis tick marks
    scale_y_discrete(labels=paste(c(as.character(Base2$Country_day)))) + 
    #average lines
    geom_vline(xintercept=mean(Base2$Heart) ,linetype="solid", color=plot_col2[4], 
               size=1.25, alpha = 0.65) +
    ggplot2::annotate(geom="text",x=mean(Base2$Heart)-0.04,y=42,label=paste(round(mean(Base2$Heart),3)),size=3.5,col=plot_col2[4])+
    theme_light() + ylab("")+ 
    #xlab("Ratio of years of life lost due to COVID-19 vs heart diseases") +
    xlab("C. Ratio of years of life lost due to COVID-19 vs heart diseases") +
    theme(panel.border = element_blank(), axis.ticks.y = element_blank()
          , axis.text.y=element_text(colour="gray20"))
#heart_base_plot
#Fig1_heart.pdf -- without xlab C
```


### Excess Panel

**Baseline plot: Excess**

```{r, echo=FALSE,eval=TRUE,warning = FALSE}
excess_base_plot <-ggplot(Base2, aes(x=Excess, y=Country)) +
    geom_point(shape=25,fill=plot_col2[5],color=plot_col2[5], size=3.5, alpha=1) +
    #relabel y-axis tick marks
    xlim(0,1) +
    scale_y_discrete(labels=paste(c(as.character(Base2$Country_day)))) + 
    #average lines
    geom_vline(xintercept=mean(Base2$Excess,na.rm=TRUE) ,linetype="solid", color=plot_col2[5], 
               size=1.25, alpha = 0.65) +
    ggplot2::annotate(geom="text",x=mean(Base2$Excess,na.rm=TRUE)+0.1,y=42,label=paste(round(mean(Base2$Excess,na.rm=TRUE),3)),size=3.5,col="goldenrod3")+ #color too light as yellow#plot_col2[5])+
    theme_light() + ylab("") + 
    #xlab("Ratio of years of life lost due to COVID-19 vs excess mortality") +
    xlab("D. Ratio of years of life lost due to COVID-19 vs excess mortality") +
    theme(panel.border = element_blank(), axis.ticks.y = element_blank()
          , axis.text.y=element_text(colour="gray20"))
#excess_base_plot
#Fig1_excess.pdf -- without xlab D
```


### Panels 1-4
**Putting together the panels Flu/Transport/Heart/Excess as 4 panels**

```{r, echo=FALSE,eval=TRUE,warning = FALSE, fig.width=12, fig.height=7}
grid.arrange(flu_base_plot,transport_base_plot,heart_base_plot,excess_base_plot, ncol=2)
```

[2x2 panel saved as `Fig1.pdf` in Figures folder on DB, 10x12.5.]{color="blue"}

[2x2 panel saved as `Fig1.eps` in Figures folder on DB, maintain aspect ratio height=900.]{color="blue"}

## Fig 1 Deaths-weighted averages in a table for SI

This LaTex table (for insertion into SI tex file) is so we can say  "The conclusions are robust to using weighted averages (SI)".

```{r, echo=FALSE,eval=TRUE,warning = FALSE, fig.width=12, fig.height=7}
#si_table1<-data.frame(Cause=c("Influenza median year","Influenza maximum year","Transport","Heart","Excess mortality"),Mean=c(mean(Base2$`Flu (med)`,na.rm=TRUE),mean(Base2$`Flu (max)`,na.rm=TRUE),mean(Base2$Transport,na.rm=TRUE),mean(Base2$Heart,na.rm=TRUE),mean(Base2$Excess,na.rm=TRUE)),Weighted_Mean=c())
matched<-match(as.character(si1$Country),as.character(Base2$Country))
Base2$Deaths<-si1$Deaths[matched]
fig1_weight<-data.frame(Ratio=c("Ratio COVID YLL to Flu (median year) YLL","Ratio COVID YLL to Flu (max year) YLL"
                               ,"Ratio COVID YLL to Transport YLL","Ratio COVID YLL to Heart disease YLL"
                               ,"Ratio COVID YLL to Excess mortality YLL")
  ,`Deaths Weighted Average`=c(sum(Base2$Deaths*Base2$`Flu (med)`/sum(Base2$Deaths))
                          ,sum(Base2$Deaths*Base2$`Flu (max)`/sum(Base2$Deaths))
                          ,sum(Base2$Deaths*Base2$Transport/sum(Base2$Deaths))
                          ,sum(Base2$Deaths*Base2$Heart/sum(Base2$Deaths))
                          ,sum(Base2$Deaths*Base2$Excess/sum(Base2$Deaths),na.rm=TRUE)) )
x<-xtable(fig1_weight, caption = "Weighted Averages for Figure 1: Ratios of COVID YLL to Other Causes of Mortality",digits=c(3))
x
```

## Figure 2:

### Age Panel

**Baseline plot: Under 55**


```{r,echo=TRUE,eval=TRUE,warning = FALSE,message=FALSE}
#Fig 2 csv data saved:
write.csv(age2,file="../Data/Final results/figure2age_data.csv")
write.csv(gender2,file="../Data/Final results/figure2gender_data.csv")

##ordered by 55
age55_base_plot<-ggplot(age2, aes(fill=Age, y=Proportion, x=Country)) + 
    geom_bar(position="stack", stat="identity") + coord_flip() +
    scale_fill_viridis(discrete = T,option = "E", alpha=0.85) +
    #Add global average values
    ggplot2::annotate(geom="text",x=rep(44,3),y=c(0.16,0.58,0.88),label=c(0.325,0.439,0.236),size=3.5,col=c("gray50","white","white")) + #from S7 Table weighted proportions for Figure 2
    xlab("") + 
    ylab("Proportion of years of life lost by age group") +
    #ylab("A.  Proportion of years of life lost by age group") +
    theme_bw() +
    ggtitle(" ") +
    theme(panel.border = element_blank(), axis.ticks.y = element_blank()
          , axis.text.y=element_text(colour="gray20")) +
    guides(fill = guide_legend(reverse=TRUE)) #-55 first
age55_base_plot
#Fig2_age - without A.
```

### Gender Panel

**Baseline plot**

```{r,echo=TRUE,eval=TRUE,warning = FALSE,message=FALSE}
tmp_mean<-mean(gender2$YLL.m_rates,na.rm=TRUE)/mean(gender2$YLL.f_rates,na.rm=TRUE)
tmp_mean_weight<-sum(gender2$Deaths/sum(gender2$Deaths,na.rm=TRUE)*gender2$YLL_rates_male_to_female,na.rm=TRUE)
## Running through examples of log-axes
gender_base_plot<-ggplot(gender2, aes(y=Country, x=Rate)) +
  geom_point(color=gender2$Rate_color, size=4) +
  geom_segment(aes(y=Country, yend=Country, x=1, xend=Rate),
              color=gender2$Rate_color)+
  geom_vline(xintercept=1,linetype="solid"
             ,color=viridis(n=1,alpha=1,begin=0.25,end=1,option="E"),size=1.25) +
  #add in global and weighted avg values
  ggplot2::annotate(geom="text",x=c(1.9,1.85,0.66,1.4),y=c(44,43,3,3)
           ,label=c(paste(round(tmp_mean,3)),paste(round(tmp_mean_weight,3)),"Women\n more\n affected","Men\n more\n affected"),size=c(rep(3.5,2),3.7,3.7),col="gray20")+ 
  #theme_bw() + xlab("B.  Ratio of male to female years of life lost") + ylab("") + 
  theme_bw() + xlab("Ratio of male to female years of life lost") + ylab("") + 
  theme(panel.border = element_blank(), axis.ticks.y = element_blank()
          , axis.text.y=element_text(colour="gray20")
        ,plot.title = element_text(face="bold",colour="gray20", size=10))
gender_natural<-gender_base_plot + 
  scale_x_continuous(lim=c(0.5,4.3),breaks = c(0,0.5,1,2,3,4), labels = c("0","0.5","1","2","3","4")) +
  theme(axis.text.x = element_text(color = rep("gray50",6)),
        axis.ticks.x = element_line(color = rep("gray90",6),
                          size = rep(.5,6)))
gender_natural
#Fig2_gender
```

**Use log10 x-axis scale**


```{r,echo=TRUE,eval=TRUE,warning = FALSE,message=FALSE}
gender_base_plot2<-ggplot(gender2, aes(y=Country, x=Rate)) +
  geom_point(color=gender2$Rate_color, size=4) +
  geom_segment(aes(y=Country, yend=Country, x=1, xend=Rate),
              color=gender2$Rate_color)+
  geom_vline(xintercept=1,linetype="solid"
             ,color=viridis(n=1,alpha=1,begin=0.25,end=1,option="E"),size=1.25) +
  #add in global and weighted avg values
  ggplot2::annotate(geom="text",x=c(1.7,1.6,0.93,1.1),y=c(44,43,4,4)
           ,label=c(paste(round(tmp_mean,3)),paste(round(tmp_mean_weight,3)),"Women\n more\n affected","Men\n more\n affected"),size=c(rep(3.5,2),3.3,3.3),col="gray20")+ 
  theme_bw() + xlab("B.  Ratio of male to female years of life lost") + ylab("") + 
  #theme_bw() + xlab("Ratio of male to female years of life lost") + ylab("") + 
  theme(panel.border = element_blank(), axis.ticks.y = element_blank()
          , axis.text.y=element_text(colour="gray20")
        ,plot.title = element_text(face="bold",colour="gray20", size=10))
gender_log<-gender_base_plot2 + scale_x_continuous(trans='log10',breaks = c(0.9,1,2,3,4)) + removeGridX()
gender_log
#SFig9_Fig2_log_gender
```

### Combined

```{r,echo=TRUE,eval=TRUE,warning = FALSE,message=FALSE}
grid.arrange(age55_base_plot,gender_natural,ncol=2)#Fig2
grid.arrange(age55_base_plot,gender_log,ncol=2) #SFig9_Fig2_log_gender
```

[Combined plot saved as `Fig2.pdf` 10x12.5.]{color="navy"}
[Combined plot with log version for gender saved as `SFig9_Fig2_log_gender.pdf` 10x12.5.]{color="navy"}
## Fig 2 Deaths-weighted averages in a table for SI

This LaTex table (for insertion into SI tex file) is so we can say  "The conclusions are robust to using weighted averages (SI)".


```{r, echo=FALSE,eval=TRUE,warning = FALSE, fig.width=12, fig.height=7}
#readRDS("../Data/Final results/ages_at_death_global.rds")
tmp<-read.csv("../Data/Final results/table_YLL_age.csv")[,-1]
countries<-as.character(sort(unique(Base$Country)))
tmp<-subset(tmp,Country%in%countries)
total_YLL<-sum(tmp$YLL_b)
total_YLL_55<-sum(tmp$YLL_b[which(tmp$Age<55)])/total_YLL
total_YLL_55_75<-sum(tmp$YLL_b[which(tmp$Age>=55&tmp$Age<75)])/total_YLL ### 0-4 --> 0, 5-9--> 5... 70-74->70
total_YLL_75<-sum(tmp$YLL_b[which(tmp$Age>=75)])/total_YLL
total_YLL_by_age<-data.frame(Age=c("Under 55","55-75","75+"),Weighted_Avg=c(total_YLL_55,total_YLL_55_75,total_YLL_75))

x<-xtable(total_YLL_by_age, caption = "Weighted Proportions for Figure 2: Proportion of total global YLL by age",digits=c(3))
x
```

[Adding groups under 55 and between 55-75, we get `r sum(total_YLL_55+total_YLL_55_75)`, or 3/4 (76%) of the total YLL comes from ages under 75.]{color="blue"}

## SI figures

### YLL and the ratio of YLL to other causes of death both grow nonlinearly/exponentially(?) with days since

**YLL grows nonlinearly with Days**

```{r,echo=TRUE,eval=TRUE,warning = FALSE,message=FALSE}
si1_base_plot <- ggplot(data=si1,aes(x=Diff.Days,y=YLL.b_abs)) + geom_point(alpha=0.5) + xlim(40,200) +
                  stat_smooth(method = 'lm', formula = y ~ poly(x,2), color=plot_col[1], se= TRUE) +
                  theme_bw()+ xlab("Days since first COVID-19 case") 
                  
si1_plot<-si1_base_plot + ylab("Years of life lost") +
  geom_text(data=si1 %>% filter(Country=="USA"|Country=="Brazil"|Country=="Algeria"|
                   Country=="Chile"|Country=="Italy"|Country=="Spain"|Country=="England"|
                   #Country=="France"|
                     Country=="Mexico"|Country=="Sweden"|Country=="Canada")
                   ,aes(label=Country),hjust=-0.0, vjust=0.0
                   ,position = position_jitter(width=0.1,height=0.1))
si1_plot
```

* Then plots "days since" against absolute YLL on log y-axis

```{r,echo=TRUE,eval=TRUE,warning = FALSE,message=FALSE}
si1_log_plot<- si1_base_plot + scale_y_continuous(trans='log10') + ylab("Log scale Years of life lost") +
                geom_text(data=si1 %>% filter(Country=="USA"|Country=="Brazil"|Country=="Algeria"|
                   Country=="Chile"|Country=="Italy"|Country=="Norway"|Country=="Romania"|
                   #Country=="France"|
                     Country=="Mexico"|Country=="Sweden"|Country=="Canada"|
                   Country=="Taiwan"|Country=="South Africa"|Country=="Japan")
                   ,aes(label=Country),hjust=-0.0, vjust=0.0
                   ,position = position_jitter(width=0.05,height=0.05))
si1_log_plot
```

* Compare the two plots, side by side
```{r,echo=TRUE,eval=TRUE,warning = FALSE,message=FALSE}
grid.arrange(si1_plot,si1_log_plot,ncol=2)
```

[Above plot saved as `Figures/SFig1.pdf` in DB.]{color="blue"}

**Ratio of YLL to other causes of death grows nonlinearly with Days**

* Plots "days since" against ratio covid YLL/flu YLL

```{r,echo=TRUE,eval=TRUE,warning = FALSE,message=FALSE}
si_alpha<-0.5
##FLU
si2_flu_base <- ggplot(data=Base2,aes(x=Diff.Days,y=`Flu (med)`)) + xlim(40,200) +
                  geom_point(color=plot_col2[3], alpha=si_alpha) +
                  stat_smooth(method = 'lm', formula = y ~ poly(x,2), color=plot_col[3], se= TRUE) +
                  theme_bw()+ xlab("Days since first COVID case") 
si2_flu<- si2_flu_base + ylab("Ratio COVID YLL/Flu YLL") + 
  ggtitle("FLU") +
  geom_text(data=Base2 %>% filter(Country=="USA"|Country=="Brazil"|
                   Country=="Chile"|Country=="Italy"|Country=="Spain"|Country=="England"|Country=="Algeria"|
                   Country=="Mexico"|Country=="Sweden"
                   |Country=="Canada")
                   ,aes(label=Country),hjust=-0.0, vjust=0.0
                   ,position = position_jitter(width=0.2,height=0.2))
si2_flu
##Transport
si2_transport_base <- ggplot(data=Base2,aes(x=Diff.Days,y=Transport)) + xlim(40,200) +
                  geom_point(color=plot_col2[1], alpha=si_alpha) +
                  stat_smooth(method = 'lm', formula = y ~ poly(x,2), color=plot_col[3], se= TRUE) +
                  theme_bw()+ xlab("Days since first COVID case") 
si2_transport<- si2_transport_base + ylab("Ratio COVID YLL/Transport YLL") + 
  ggtitle("TRANSPORT") +
  geom_text(data=Base2 %>% filter(Country=="USA"|
                   Country=="Chile"|Country=="Italy"|Country=="Algeria"|
                   Country=="Mexico"|Country=="Sweden"|Country=="Canada"|
                   Country=="Taiwan"|Country=="South Africa"|Country=="Japan")
                   ,aes(label=Country),hjust=-0.0, vjust=0.0
                   ,position = position_jitter(width=0.2,height=0.2))
si2_transport
##Heart
si2_heart_base <- ggplot(data=Base2,aes(x=Diff.Days,y=Heart)) + xlim(40,200) +
                  geom_point(color=plot_col2[4], alpha=si_alpha) +
                  stat_smooth(method = 'lm', formula = y ~ poly(x,2), color=plot_col[3], se= TRUE) +
                  theme_bw()+ xlab("Days since first COVID case") 
si2_heart<- si2_heart_base + ylab("Ratio COVID YLL/Heart disease YLL") + 
  ggtitle("HEART DISEASE") +
  geom_text(data=Base2 %>% filter(Country=="USA"|Country=="Philippines"|
                   Country=="Chile"|Country=="Italy"|#Country=="New Zealand"|
                   Country=="Mexico"|Country=="Canada"|
                   Country=="Taiwan"|Country=="South Africa"|Country=="Japan")
                   ,aes(label=Country),hjust=-0.0, vjust=0.0
                   ,position = position_jitter(width=0.01,height=0.01))
si2_heart
##Excess
si_excess_data<-Base2[which(!is.na(Base2$Excess)),]
si_excess_data$Country<-factor(as.character(si_excess_data$Country))
si2_excess_base <- ggplot(data=si_excess_data,aes(x=Diff.Days,y=Excess)) + xlim(55,210) +
                  geom_point(color=plot_col2[5], alpha=1) +
                  stat_smooth(method = 'lm', formula = y ~ poly(x,2), color=plot_col[3], se= TRUE) +
                  #stat_smooth(method = "loess", formula = y ~ x, size = 1) +
                  theme_bw()+ xlab("Days since first COVID case") 
si2_excess<- si2_excess_base + ylab("Ratio COVID YLL/Excess mortality YLL") + 
  ggtitle("EXCESS MORTALITY") +
  geom_text(data=si_excess_data %>% filter(Country!="Sweden")
                   ,aes(label=Country),hjust=-0.0, vjust=0.0
                   ,position = position_jitter(width=0.01,height=0.01))
si2_excess
```

* Then plots "days since" against ratio covid YLL/flu YLL on log y-axis

```{r,echo=TRUE,eval=TRUE,warning = FALSE,message=FALSE}
##FLU
si2_logflu<- si2_flu_base + scale_y_continuous(trans='log10') + ylab("Log scale ratio COVID YLL/Flu YLL") +
                ggtitle("FLU") +
                geom_text(data=Base2 %>% filter(Country=="USA"|Country=="Brazil"|
                   Country=="Chile"|Country=="Italy"|Country=="Norway"|Country=="Romania"|Country=="Algeria"|
                   Country=="Portugal"|Country=="Mexico"#|Country=="Sweden"
                   |Country=="Canada"|
                   Country=="Taiwan"|Country=="South Africa"|Country=="Japan")
                   ,aes(label=Country),hjust=-0.0, vjust=0.0
                   ,position = position_jitter(width=0.1,height=0.1))
si2_logflu
##Transport
si2_logtransport<- si2_transport_base + scale_y_continuous(trans='log10') + ylab("Log scale ratio COVID YLL/Transport YLL") +
                ggtitle("TRANSPORT") +
                geom_text(data=Base2 %>% filter(Country=="USA"|Country=="New Zealand"|#|Country=="Algeria"
                   Country=="Chile"|Country=="Italy"|
                   Country=="Mexico"|Country=="Sweden"|Country=="Canada"|
                   Country=="Taiwan"|Country=="South Africa"|Country=="Japan")
                   ,aes(label=Country),hjust=-0.0, vjust=0.0
                   ,position = position_jitter(width=0.1,height=0.1))
si2_logtransport
##Heart
si2_logheart<- si2_heart_base + scale_y_continuous(trans='log10') + ylab("Log scale ratio COVID YLL/Heart YLL") +
                ggtitle("HEART DISEASE") +
                geom_text(data=Base2 %>% filter(Country=="USA"|
                   Country=="Chile"|Country=="Italy"|
                   #Country=="Mexico"|
                     Country=="New Zealand"|
                     Country=="Canada"|
                   Country=="Taiwan"|Country=="South Africa"|Country=="Japan")
                   ,aes(label=Country),hjust=-0.0, vjust=0.0
                   ,position = position_jitter(width=0.1,height=0.1))
si2_logheart
##Excess
si2_logexcess<- si2_excess_base + scale_y_continuous(trans='log10') + ylab("Log scale ratio COVID YLL/Excess mortality YLL") +
                ggtitle("EXCESS MORTALITY") +
                geom_text(data=si_excess_data %>% filter(Country!="Netherlands" & Country!="Italy" & Country !="Sweden")
                   ,aes(label=Country),hjust=-0.0, vjust=0.0
                   ,position = position_jitter(width=0.1,height=0.1))
si2_logexcess
```

* Compare the two plots, side by side
```{r, echo=FALSE,eval=TRUE,warning = FALSE, fig.width=13, fig.height=6}
##Flu
grid.arrange(si2_flu,si2_logflu,ncol=2)#save 7x12.5
##Transport
grid.arrange(si2_transport,si2_logtransport,ncol=2)#save 7x12
##Heart
grid.arrange(si2_heart,si2_logheart,ncol=2)#save 7x12
##Excess
grid.arrange(si2_excess,si2_logexcess,ncol=2)#save 6x10

#grid.arrange(si2_flu,si2_logflu,si2_transport,si2_logtransport,si2_heart,si2_logheart,si2_excess,si2_logexcess,ncol=2) #save 12x12
```

[Above plots saved as: full version `Figures/SFig2_all.pdf`; then 2 paneled versions by cause  `Figures/SFig3_flu.pdf`,`Figures/SFig4_transport.pdf`,`Figures/SFig5_heart.pdf`,`Figures/SFig6_excess.pdf` in DB.]{color="blue"}

### F2-A SI plots

```{r,echo=FALSE,eval=TRUE,warning = FALSE,message=FALSE}
age2_55<-subset(age2,Age=="Under 55"&(Country!="GLOBAL AVERAGE"&Country!=""))
matched<-match(as.character(age2_55$Country),as.character(days$Country))
age2_55$Diff.Days<-days$Diff.Days[matched[which(!is.na(matched))]]
age2_75<-subset(age2,Age=="75+"&(Country!="GLOBAL AVERAGE"&Country!=""))
matched<-match(as.character(age2_75$Country),as.character(days$Country))
age2_75$Diff.Days<-days$Diff.Days[matched[which(!is.na(matched))]]
```

**Less than 55 against Days**

```{r,echo=TRUE,eval=TRUE,warning = FALSE,message=FALSE}
si3_55_base<-ggplot(data=age2_55,aes(x=Diff.Days,y=Proportion)) + 
                  geom_point(color=plot_col[1], alpha=si_alpha) +
                  stat_smooth(method = 'lm', formula = y ~ poly(x,2), color=plot_col[3], se= TRUE) +
                  theme_bw()+ xlab("Days since first COVID-19 case") 
si3_55 <- si3_55_base + ylab("Proportion of YLL for Age under 55") + ggtitle("Age under 55") +
  geom_text(data=age2_55 %>% filter(Country=="USA"|Country=="Brazil"|
                   Country=="Chile"|Country=="Italy"|Country=="Spain"|Country=="England"|
                   Country=="France"|Country=="Bangladesh"|Country=="South Africa"|Country=="Canada")
                   ,aes(label=Country),hjust=-0.0, vjust=0.0
                   ,position = position_jitter(width=0.01,height=0.01))
si3_55
```

**75+ against Days**

```{r,echo=TRUE,eval=TRUE,warning = FALSE,message=FALSE}
si3_75_base<-ggplot(data=age2_75,aes(x=Diff.Days,y=Proportion)) + 
                  geom_point(color=plot_col[3], alpha=si_alpha) +
                  stat_smooth(method = 'lm', formula = y ~ poly(x,2), color=plot_col[3], se= TRUE) +
                  theme_bw()+ xlab("Days since first COVID-19 case") 
si3_75 <- si3_75_base + ylab("Proportion of YLL for Age over 75") + ggtitle("Age over 75") +
  geom_text(data=age2_75 %>% filter(Country=="USA"|Country=="Brazil"|
                   Country=="Chile"|Country=="Italy"|Country=="Spain"|Country=="England"|
                   Country=="France"|Country=="Bangladesh"|Country=="South Africa"|Country=="Canada")
                   ,aes(label=Country),hjust=-0.0, vjust=0.0
                   ,position = position_jitter(width=0.01,height=0.01))
si3_75
```


Age distribution of YLL is not strongly influenced by the duration since the outbreak (SI Figure 3): 

```{r,echo=TRUE,eval=TRUE,warning = FALSE,message=FALSE}
legend_val<-c("Under 55" = plot_col[3], "75+" = plot_col[1])
si3_base <-ggplot(data=age2_55,aes(x=Diff.Days,y=Proportion,group = 1)) + 
                  geom_point(color=plot_col[1], alpha=si_alpha) +
                  stat_smooth(data=age2_55,aes(x=Diff.Days,y=Proportion,color="Under 55"),method = 'lm'
                              , formula = y ~ poly(x,2),  se= TRUE) +
                  geom_point(data=age2_75,aes(x=Diff.Days,y=Proportion),color=plot_col[3], alpha=si_alpha) +
                  stat_smooth(data=age2_75,aes(x=Diff.Days,y=Proportion, color="75+"),method = 'lm'
                              , formula = y ~ poly(x,2), se= TRUE) +
                  scale_colour_manual(values = legend_val) +
                  theme_bw()+ labs(x="Days since first COVID-19 case",y="Proportion of YLL",color="Legend") +
                  theme(panel.border = element_blank(), axis.ticks.y = element_blank()
                  ,legend.position = "top",legend.justification = c(0, 1)
                  , axis.text.y=element_text(colour="gray20")
                  ,plot.title = element_text(face="bold",colour="gray20", size=10))
si3_base
```

[Combined plot of 75+ and <55 saved as `Figures/SFig7.pdf` in DB. ]{color="blue"}

## F2-B SI plots

```{r,echo=FALSE,eval=TRUE,warning = FALSE,message=FALSE}
gender3<-subset(gender2,Country!="WEIGHTED AVERAGE"&Country!="AVERAGE"&Country!="")
matched<-match(as.character(gender3$Country),as.character(days$Country))
gender3$Diff.Days<-days$Diff.Days[matched[which(!is.na(matched))]]
```

**Male YLL against days**
```{r,echo=TRUE,eval=TRUE,warning = FALSE,message=FALSE}
si4_male_base<-ggplot(data=gender3,aes(x=Diff.Days,y=YLL.m_rates)) + 
                  geom_point(color=plot_col[4], alpha=si_alpha) +
                  stat_smooth(method = 'lm', formula = y ~ poly(x,2), color=plot_col[3], se= TRUE) +
                  theme_bw()+ xlab("Days since first COVID-19 case") 
si4_male <- si4_male_base + ylab("YLL rates for men") + ggtitle("Male YLL rates over time") +
  geom_text(data=gender3 %>%filter(Country=="USA"|Country=="Belgium"|
                   Country=="Spain"|Country=="Italy"|Country=="Switzerland"|Country=="Taiwan"|
                   Country=="Japan"|Country=="Pakistan"|Country=="Portugal"|Country=="Bolivia")
                   ,aes(label=Country),hjust=-0.0, vjust=0.0
                   ,position = position_jitter(width=0.01,height=0.01))
si4_male
```

**Female YLL against days**
```{r,echo=TRUE,eval=TRUE,warning = FALSE,message=FALSE}
si4_female_base<-ggplot(data=gender3,aes(x=Diff.Days,y=YLL.f_rates)) + 
                  geom_point(color=plot_col[4], alpha=si_alpha) +
                  stat_smooth(method = 'lm', formula = y ~ poly(x,2), color=plot_col[3], se= TRUE) +
                  theme_bw()+ xlab("Days since first COVID-19 case") 
si4_female <- si4_female_base + ylab("YLL rates for women") + ggtitle("Female YLL rates over time") +
  geom_text(data=gender3 %>% filter(Country=="USA"|Country=="Belgium"|
                   Country=="Spain"|Country=="Italy"|Country=="Switzerland"|Country=="Taiwan"|
                   Country=="Japan"|Country=="Pakistan"|Country=="Portugal"|Country=="Bolivia")
                   ,aes(label=Country),hjust=-0.0, vjust=0.0
                   ,position = position_jitter(width=0.02,height=0.02))
si4_female
```

gender distribution of YLL not strongly influenced by the duration since the outbreak (SI Figure 4):

```{r,echo=TRUE,eval=TRUE,warning = FALSE,message=FALSE}
legend_val<-c("Male" = plot_col[3], "Female" = plot_col[1])
si4<-ggplot(data=gender3,aes(x=Diff.Days,y=YLL.m_rates)) + 
                  geom_point(color=plot_col[3], alpha=si_alpha) +
                  stat_smooth(aes(x=Diff.Days,y=YLL.m_rates,color="Male"),method = 'lm', formula = y ~ poly(x,2), se=TRUE) +
                  geom_point(aes(x=Diff.Days,y=YLL.f_rates),color=plot_col[1], alpha=si_alpha) +
                  stat_smooth(aes(x=Diff.Days,y=YLL.f_rates,color="Female"), method = 'lm', formula = y ~ poly(x,2),se=TRUE) +
                  theme_bw()+ labs(x="Days since first COVID-19 case",y="YLL Rates",color="Legend") +
                  scale_colour_manual(values = legend_val) +
                  theme(panel.border = element_blank(), axis.ticks.y = element_blank()
                  ,legend.position = "top",legend.justification = c(0, 1)
                  , axis.text.y=element_text(colour="gray20")
                  ,plot.title = element_text(face="bold",colour="gray20", size=10))
si4
```

[Combined plot saved as `Figures/SFig8.pdf` in DB.]{color="blue"}



# Notes

551,384 global covid deaths (https://coronavirus.jhu.edu/map.html) accessed July 9, 2020.

Covered deaths in the dataset: 383,028. Proportion `r (383028/551384)`